---
title: "Primary Analyses"
author: "Lucy King"
output: html_document
---

```{r}
#Libraries
library(tidyverse)
library(haven)
library(lubridate)
library(modelr)

#Files 
hormones_file <- "~/Desktop/ELS/cort_dhea/data/data_final/hormones_final_20190506.csv"
els_file <- "~/Desktop/ELS/stress_data/data/els_severity_scores_20190221.csv"
tanner_t1_file <- "~/Desktop/ELS/cort_dhea/data/data_final/tanner.T1.sav"
tanner_t2_s1_file <- "~/Desktop/ELS/cort_dhea/data/data_final/tanner_s1.T2.sav"
tanner_t2_s2_file <- "~/Desktop/ELS/cort_dhea/data/data_final/tanner_s2.T2.sav"
demo_t1_file <- "~/Desktop/ELS/cort_dhea/data/data_final/demographics_golden_standard.T1.sav"
demo_t2_file <- "~/Desktop/ELS/cort_dhea/data/data_final/demographics_golden_standard.T2.sav"
birthdate_file <- "~/Desktop/ELS/cort_dhea/data/data_final/ELS T1 Birthdates.sav"

#Functions
source("~/Desktop/ELS/cort_dhea/cort_dhea_sync/els_cort_dhea/winsorize.R")
source("~/Desktop/ELS/cort_dhea/cort_dhea_sync/els_cort_dhea/identify_outliers_histogram.R")
```

# Read in data

## Hormone data

```{r}
hormones <- 
  read_csv(hormones_file) %>% 
  filter(timepoint == "T1" | timepoint == "T2") %>% 
  select(
    ELS_ID:birth_control,
    medication_binary,
    testosterone_mean,
    cortisol_mean,
    dhea_mean,
    testosterone_mean_clean,
    cortisol_mean_clean,
    dhea_mean_clean
  ) %>% 
  mutate(
    date_collection = if_else(
      is.na(date_collection),
      date(scan_date), date(date_collection)
    ),
    date_collection = if_else( #using 85's session date (missing scan date and hormone date)
      ELS_ID == 85 & timepoint == "T2",
      date("2016-10-16"), date_collection
    ),
    medication_binary = as.factor(medication_binary)
  )

hormones_t1 <-
  hormones %>% 
  filter(timepoint == "T1")

hormones_t2 <-
  hormones %>% 
  filter(timepoint == "T2")
```


## Tanner 

### Tanner T1
There are n = 14 at T1 who are not in early puberty. Removing these participants.
```{r}
##T1
tanner_t1 <-
  read_sav(tanner_t1_file) %>% 
  rename_all(
    funs(str_replace(., ".T1", ""))
  ) %>% 
  rename(timepoint = Timepoint) %>% 
  select(
    -Subject_ID_Timepoint,
    -Session_Type
  ) %>% 
  mutate(
    early_puberty = 
      (tanner_female_pubhair < 4 & tanner_female_breasts < 4) |
      (tanner_male_pubhair < 4 & tanner_male_testes < 4)
  )
    
tanner_t1 %>% 
  distinct(ELS_ID, early_puberty) %>% 
  count(early_puberty)

keep_IDs <- 
  tanner_t1 %>% 
  distinct(ELS_ID, early_puberty) %>% 
  filter(early_puberty == TRUE) %>% 
  pull(ELS_ID)

exclude_IDs <- 
  tanner_t1 %>% 
  filter(is.na(early_puberty)) %>% 
  pull(ELS_ID)
```

### counts for hormones and why missing
```{r}
included_T1_file <- "~/Desktop/ELS/cort_dhea/data/data_final/included_T1.csv"
included_T1 <- read_csv(included_T1_file)

hormones_earlypuberty <-
  hormones %>% 
  filter(ELS_ID %in% keep_IDs)

## T1 numbers______________________________________________________
hormones_earlypuberty %>% 
  filter(timepoint == "T1") %>% 
  filter(!is.na(dhea_mean_clean) | !is.na(cortisol_mean_clean) | !is.na(testosterone_mean_clean)) %>% 
  ungroup() %>% 
  summarise(n_distinct(ELS_ID))

hormones_earlypuberty %>% 
  filter(timepoint == "T1") %>% 
  filter(!is.na(dhea_mean_clean)) %>% 
  ungroup() %>% 
  summarise(n_distinct(ELS_ID))

hormones_earlypuberty %>% 
  filter(timepoint == "T1") %>% 
  filter(!is.na(cortisol_mean_clean)) %>% 
  ungroup() %>% 
  summarise(n_distinct(ELS_ID))

hormones_earlypuberty %>% 
  filter(timepoint == "T1") %>% 
  filter(!is.na(testosterone_mean_clean)) %>% 
  ungroup() %>% 
  summarise(n_distinct(ELS_ID))

## T2 numbers______________________________________________________

hormones_earlypuberty %>% 
  filter(timepoint == "T2") %>% 
  filter(!is.na(dhea_mean_clean) | !is.na(cortisol_mean_clean) | !is.na(testosterone_mean_clean)) %>% 
  ungroup() %>% 
  summarise(n_distinct(ELS_ID))

hormones_earlypuberty %>% 
  filter(timepoint == "T2") %>% 
  filter(!is.na(cortisol_mean_clean)) %>% 
  ungroup() %>% 
  summarise(n_distinct(ELS_ID))

hormones_earlypuberty %>% 
  filter(timepoint == "T2") %>% 
  filter(!is.na(dhea_mean_clean)) %>% 
  ungroup() %>% 
  summarise(n_distinct(ELS_ID))

hormones_earlypuberty %>% 
  filter(timepoint == "T2") %>% 
  filter(!is.na(testosterone_mean_clean)) %>% 
  ungroup() %>% 
  summarise(n_distinct(ELS_ID))


## cortisol OR DHEA at T1 OR T2

hormones_earlypuberty %>% 
  filter(timepoint == "T2" | timepoint == "T1") %>% 
  filter(!is.na(dhea_mean_clean) | !is.na(cortisol_mean_clean) | !is.na(testosterone_mean_clean)) %>% 
  ungroup() %>% 
  summarise(n_distinct(ELS_ID))

## cortisol at T1 OR T2
hormones_earlypuberty %>% 
  filter(timepoint == "T2" | timepoint == "T1") %>% 
  filter(!is.na(cortisol_mean_clean)) %>% 
  ungroup() %>% 
  summarise(n_distinct(ELS_ID))
```

N = 214 participants at T1
N = 14 in later puberty = 200 who meet criteria

T1:

N = 26/200 did not complete any hormone samples at T1; 3/200 qns for cort, dhea, & testosterone at T1 = 171 at T1
N = 19 additional qns cortisol at T1; 1 mean below limit at T1 = 151 cortisol at T1
N = 1 mean below limit dhea at T1 = 170 at T1
N = 3 additional qns testosterone at T1 (additional 3) = 168 at T1

T2:

N = 4/200 did not complete hormone samples, 13/200 dropped out at T2, 2/200 opted out, 15/200 timed out, 21/200 not yet returned,  2/200 qns dhea, cortisol, and testosterone = 143 at T2
N = 6/140 additional qns cortisol = 136 cortisol at T2
N = 3/140 additional qns dhea = 140 dhea at T2
N = 0/140 qns additional testosterone = 143 dhea at T2

cortisl OR DHEA OR testosterone at either T1 OR T2:

N = 22/200 did not complete dhea OR cortisol OR testosterone at T1 OR T2 = 178 analyzable 
1/22 did not complete hormones samples at T1 or T2
6/22 did not complete at T1; dropped out at T2
6/22 did not complete at T1; timed out at T2
7/22 did not complete at T1; not yet returned T2
1/25 qns at T1; timed out at T2
1/25 qns at T1 and T2


### Tanner T2
*Given multiple sessions at T2, identify tanner scores associated with date of hormone collection. If no exact date match, use closest date.* 
```{r}
clean_tanner <- function(tan_s1T2, tan_s2T2, closest_tanner) {
  case_when(
    is.na(tan_s1T2) ~ tan_s2T2,
    is.na(tan_s2T2) ~ tan_s1T2,
    closest_tanner == "s1" ~ tan_s1T2,
    closest_tanner == "s2" ~ tan_s2T2
  )
}

tanner_t2 <-
  read_sav(tanner_t2_s1_file) %>% 
  left_join(read_sav(tanner_t2_s2_file), by = "ELS_ID") %>% 
  left_join(
    hormones_t2 %>% 
      select(ELS_ID, date_collection),
    by = "ELS_ID"
  ) %>% 
  mutate_at(
    vars(tanner_date_s1.T2, tanner_date_s2.T2),
    funs(parse_date_time(., "Ymd"))
  ) %>% 
  mutate(
    date_match = case_when(
      date_collection == tanner_date_s1.T2 ~ 1,
      date_collection == tanner_date_s2.T2 ~ 1,
      TRUE ~ 0
    ),
    s1_hd_diff = date_collection %--% tanner_date_s1.T2 / days(1),
    s2_hd_diff = date_collection %--% tanner_date_s2.T2 / days(1)
  ) %>% 
  filter(
    date_match == 1 | 
      (s1_hd_diff <= 90 & s1_hd_diff >= -90) |
      (s2_hd_diff <= 90 & s2_hd_diff >= -90)
  ) %>% 
  mutate(
    closest_tanner = case_when(
      abs(s1_hd_diff) < abs(s2_hd_diff) ~ "s1",
      abs(s2_hd_diff) < abs(s1_hd_diff) ~ "s2",
      s1_hd_diff == 0 ~ "s1",
      abs(s1_hd_diff) == abs(s2_hd_diff) ~ "s1",
      is.na(s2_hd_diff) ~ "s1",
      is.na(s1_hd_diff) ~ "s2",
      TRUE ~ NA_character_
    ),
    tanner_female_pubhair = clean_tanner(
      tanner_female_pubhair_s1.T2,
      tanner_female_pubhair_s2.T2,
      closest_tanner
    ),
    tanner_female_breasts = clean_tanner(
      tanner_female_breasts_s1.T2,
      tanner_female_breasts_s2.T2,
      closest_tanner
    ),
    tanner_male_pubhair = clean_tanner(
      tanner_male_pubhair_s1.T2,
      tanner_male_pubhair_s2.T2,
      closest_tanner
    ),
    tanner_male_testes = clean_tanner(
      tanner_male_testes_s1.T2,
      tanner_male_testes_s2.T2,
      closest_tanner 
    )
  ) %>% 
  rename(
    timepoint = Timepoint.x 
  ) %>% 
  select(
    ELS_ID,
    timepoint,
    tanner_female_pubhair:tanner_male_testes
  )
```
```{r warning=FALSE}
tanner <-
  tanner_t1 %>% 
  bind_rows(tanner_t2) %>% 
  group_by(ELS_ID) %>% 
  mutate(
    tanner_pb = if_else(
      is.na(tanner_female_pubhair), 
      tanner_male_pubhair, 
      tanner_female_pubhair
    ),
    tanner_bt = if_else(
      is.na(tanner_female_breasts),
      tanner_male_testes,
      tanner_female_breasts
    )
  ) %>% 
  ungroup() %>% 
  mutate(
    tanner_av = pmap_dbl(
      select(., c("tanner_pb", "tanner_bt")),
      function(...) mean(c(...), na.rm = TRUE)
    )
  )

```

## Demographics
```{r}
demo <-
  read_sav(demo_t1_file) %>% 
  mutate(
    ELS_ID = as.integer((ELS_ID)),
    session_type = "first session"
  ) %>% 
  select(
    ELS_ID,
    session_type,
    timepoint = Timepoint,
    bmi = T1_BMI,
    race_pr = T1_KSADS_Child_Race_by_P,
    eth_pr = T1_Qualtrics_Ethnicity
  ) %>% 
  bind_rows(
    read_sav(demo_t2_file) %>% 
      mutate(
        ELS_ID = as.integer((ELS_ID)),
        session_type = case_when(
          str_detect(T2_Session_Type, "\\.2") == TRUE ~ "second session",
          str_detect(T2_Session_Type, "\\.5") == TRUE ~ "session after braces",
          TRUE ~ "first session"
        )
      ) %>% 
      select(
        ELS_ID,
        session_type,
        timepoint = Timepoint,
        bmi = T2_BMI
      ) 
  ) %>% 
  left_join(
    read_sav(demo_t1_file) %>% 
      select(
        ELS_ID,
        parent_ed_t1 = T1_Parent_Education,
        inr_t1 = T1_Demo_Income_To_Needs
      ) %>% 
      mutate(ELS_ID = as.integer(ELS_ID)),
    by = "ELS_ID"
  ) %>% 
  left_join(
    read_sav(birthdate_file) %>% 
      select(
        ELS_ID,
        birth_date = T1_Child_Birthdate
      ) %>% 
      mutate(ELS_ID = as.integer(ELS_ID)),
    by = "ELS_ID"
  )
```


# Join data frames
```{r}
cd <-
  hormones %>% 
  left_join(
    read_csv(els_file) %>% 
      select(ELS_ID, sumsev_threat_t1, sumsev_threat_t2, sumsev_type_t1, sumsev_type_t2),
    by = "ELS_ID"
  ) %>% 
  left_join(demo, by = c("ELS_ID", "timepoint", "session_type")) %>% 
  left_join(tanner, by = c("ELS_ID", "timepoint")) %>% 
  filter(ELS_ID < 300, ELS_ID %in% keep_IDs) %>% 
  mutate(
    inr_t1 = if_else(
      inr_t1 == 999 | inr_t1 == 888,
      NA_real_, inr_t1
    ),
    birth_date = parse_date_time(birth_date, "Ymd"),
    child_age =  case_when(
      !is.na(date_collection) ~ birth_date %--%  date_collection / years(1),
      is.na(date_collection) ~ birth_date %--%  scan_date / years(1),
      TRUE ~ NA_real_
    )
  ) %>% 
  distinct(ELS_ID, date_collection, .keep_all = TRUE) %>% 
  select(
    -tanner_female_pubhair:-tanner_male_testes,
    -early_puberty
  ) %>% 
  select(
    ELS_ID:birth_control,
    medication_binary,
    cort_raw = cortisol_mean,
    dhea_raw = dhea_mean,
    test_raw = testosterone_mean,
    cort_clean = cortisol_mean_clean,
    dhea_clean = dhea_mean_clean,
    test_clean = testosterone_mean_clean,
    sumsev_threat_t1:child_age
  ) %>% 
  arrange(ELS_ID, timepoint, date_collection) %>% 
  group_by(ELS_ID, timepoint) %>% 
  mutate(
    n_session = 1:n()
  ) %>% 
  # if more than 1 session, use first
  filter(n_session == 1) %>% 
  ungroup()
```
# Calculate interval between T1 and T2

```{r}
interval <-
  cd %>% 
  select(
    ELS_ID,
    timepoint,
    child_age,
    tanner_av,
    cort_clean,
    dhea_clean,
    test_clean
  ) %>% 
  gather(key, variable, child_age:test_clean) %>% 
  unite(new, timepoint, key) %>% 
  spread(new, variable) %>% 
  mutate(
    interval_yr = T2_child_age - T1_child_age
  ) 
```

## Distributions of interval
```{r}
identify_outliers_hist(interval, interval_yr)
```

## Remove outliers for interval_yr 

```{r}
interval <-
  interval %>% 
  filter(interval_yr < 3)
```


```{r}
cd <-
  cd %>% 
  left_join(
    interval %>% 
      select(ELS_ID, interval_yr, T1_cort_clean, T1_dhea_clean, T1_test_clean), 
    by = "ELS_ID"
  ) %>% 
  mutate(
    child_age_0 = if_else(
      timepoint == "T1",
      0, interval_yr
    )
  )

```


# Join T1 covariates with long-form file
```{r}
cd_t1 <-
  cd %>% 
  filter(timepoint == "T1")

cd <-
  cd %>% 
  left_join(
    cd_t1 %>% 
      select(
        ELS_ID,
        child_age_t1 = child_age,
        tanner_av_t1 = tanner_av,
        bmi_t1 = bmi
      ),
    by = "ELS_ID"
  )
```

# Wide form file

```{r}
cd_t2 <-
  cd %>% 
  filter(timepoint == "T2")

cd_wf <- 
  cd_t1 %>% 
  select(-interval_yr) %>% 
  rename(
    dhea_raw_t1 = dhea_raw,
    cort_raw_t1 = cort_raw,
    test_raw_t1 = test_raw,
    dhea_clean_t1 = dhea_clean,
    cort_clean_t1 = cort_clean,
    test_clean_t1 = test_clean,
    tanner_t1 = tanner_av,
    child_age_t1 = child_age,
    bmi_t1 = bmi,
    time_collection_t1 = time_collection,
    medication_t1 = medication_binary,
    corticosteroid_t1 = corticosteroid
  ) %>% 
  left_join(
    cd_t2 %>% 
      select(
        ELS_ID,
        interval_yr,
        dhea_raw_t2 = dhea_raw,
        cort_raw_t2 = cort_raw,
        test_raw_t2 = test_raw,
        dhea_clean_t2 = dhea_clean,
        cort_clean_t2 = cort_clean,
        test_clean_t2 = test_clean,
        tanner_t2 = tanner_av,
        child_age_t2 = child_age,
        bmi_t2 = bmi,
        time_collection_t2 = time_collection,
        medication_t2 = medication_binary,
        corticosteroid_t2 = corticosteroid
      ),
    by = "ELS_ID"
  ) %>% 
  mutate(
    cort_slope_yr = (cort_clean_t2 - cort_clean_t1) / interval_yr,
    dhea_slope_yr = (dhea_clean_t2 - dhea_clean_t1) / interval_yr,
    test_slope_yr = (test_clean_t2 - test_clean_t1) / interval_yr
  )

```

## Distributions of slope
```{r}
identify_outliers_hist(cd_wf, cort_slope_yr)

identify_outliers_hist(cd_wf, dhea_slope_yr)

identify_outliers_hist(cd_wf, test_slope_yr)
```

## Winsorize outlying dhea_slope values

```{r}
cd_wf <-
  cd_wf %>% 
  mutate(
    dhea_slope_yr_win = winsorize(dhea_slope_yr, 3),
    test_slope_yr_win = winsorize(test_slope_yr, 3)
  )
```

## Final clean data sets
```{r}
cd_comp <-
  cd %>% 
  filter(!is.na(cort_clean) | !is.na(dhea_clean) | !is.na(test_clean)) 

cd_wf_comp <-
  cd_wf %>% 
  filter(
    !is.na(cort_clean_t1) | 
      !is.na(dhea_clean_t1) | 
      !is.na(cort_clean_t2) | 
      !is.na(dhea_clean_t2) |
      !is.na(test_clean_t1) |
      !is.na(test_clean_t2) 
  )
```

```{r}
write_csv(demo, "~/Desktop/ELS/cort_dhea/data/data_final/demo.csv")
write_csv(tanner, "~/Desktop/ELS/cort_dhea/data/data_final/tanner_t1t2.csv")
write_csv(cd, "~/Desktop/ELS/cort_dhea/data/data_final/cort_dhea_all_20190509.csv")
write_csv(cd_comp, "~/Desktop/ELS/cort_dhea/data/data_final/cort_dhea_final_20190509.csv")
write_csv(cd_wf, "~/Desktop/ELS/cort_dhea/data/data_final/cort_dhea_all_wf_20190509.csv")
write_csv(cd_wf_comp, "~/Desktop/ELS/cort_dhea/data/data_final/cort_dhea_final_wf_20190509.csv")
```


